### 2024-05-23 notes/grammar-as-separation-logic.tex
Repo reorg

\documentclass[12pt]{article}
%AMS-TeX packages
\usepackage{tikz}
\usepackage[utf8]{inputenc}
\usepackage{amssymb,amsmath,amsthm,stmaryrd, wasysym} 
%geometry (sets margin) and other useful packages
\usepackage[margin=1.25in]{geometry}
\usepackage{graphicx,ctable,booktabs}
\usepackage{tikz-cd}
\usepackage{mathpartir}

\usepackage[sort&compress,square,comma,authoryear]{natbib}
\bibliographystyle{plainnat}

\newcommand{\leftmultimap}{\mathop{\rotatebox[origin=c]{180}{$\multimap$}}}
\newcommand{\Set}{\textrm{Set}}
\newcommand{\FinSet}{\mathbb F}
\newcommand{\Cat}{\textrm{Cat}}
\newcommand{\Lang}{\textrm{Lang}}
\newcommand{\Grammar}{\textrm{Grammar}}
\newcommand{\Action}{\textrm{Action}}
\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}

\newcommand\rsepimp{\mathbin{-\mkern-6mu*}}
\newcommand\lsepimp{\mathbin{\rotatebox[origin=c]{180}{$-\mkern-6mu*$}}}
\newcommand\sepconj{\mathbin{*}}
\newcommand\bunch{\mathcal W}

\begin{document}
\title{Grammars as Separation Logic Programs}
\author{Max S. New}
\maketitle

\section{Formal Languages and Formal Grammars}

Formal language theory is one of the great success stories of
theoretical computer science as a discipline, providing the foundation
for parser generators etc. The fundamental notion of formal language
theory is the formal language. Given an alphabet $\mathcal A$, a
formal language is a subset of the set of strings $L \subseteq
\mathcal A^*$. The beauty of this mathematical definition is that this
is a universal notion of language: it provides a simple specification
that can be used as a single semantic domain for many disparate
grammar formalisms without a priori limiting the languages to be
recognizable or efficiently recognizable. While formal languages
provide a simple semantic domain, they have limitations as
specifications for \emph{parsers}, one of the most ubiquitous sorts of
programs. The reason is that fundamentally a formal language $L$,
being a subset, can only properly serve by itself as the specification
for a \emph{recognizer} which takes a string and outputs a boolean. A
correct recognizer with respect to a language $L$ is a computable
version of the indicator function for the subset. But parsers output
far richer data than merely single bits, they output some kind of
\emph{parse tree}, or in reality when combined with semantic actions
produce an output of an arbitrary data type. Further, formal grammar
formalisms themselves have a richer semantics than merely a formal
language: for instance, formal grammars may be \emph{ambiguous}
meaning there is more than one parse tree for a single input string,
something which cannot be captured in a subset semantics.

This brings us to our first definition:
\begin{definition}
  A grammar between strings over a type $\mathcal A$ with semantic
  values in a type $X$ is a function
  \[ \mathcal A^* \to X \to \textrm{Type} \]
\end{definition}
That is, fundamentally a grammar is simply a constructive relation
between strings and ``things''\footnote{thanks Wadler}. Given such a
grammar $G$, a parse tree for a string $w$ with semantic meaning $x$
is an element $G\,w\,x$. We will show that just as formal languages
provide a unified semantics for languages, this notion of grammar
provides a unified semantics for grammars.

Given a grammar, we can immediately specify total and partial parsers for the grammar:
\begin{definition}
  Let $G \in \Grammar\,{\mathcal A}\,X$. A \emph{total parser} for $G$ is a function
  \[ (w:\mathcal A^*) \to (x: X) \times G\,w\,x\]
  and a (decidably) \emph{partial parser} is a function
  \[ (w : \mathcal A^*) \to 1 + ((x: X) \times G\,w\,x)\]
\end{definition}
That is, a parser outputs a semantic value and a parse tree that
``proves'' that the semantic value is represented by the
string\footnote{in a practical implementation, generating the parse
tree at runtime is unnecessary. This could possibly be modeled using
modalities for \emph{runtime irrelevance}.}. We say a grammar is
\emph{unambiguous for parsing} if for every $w$, the type
\[ (x: X) \times G\,w\,x \]
is a sub-singleton: it has at most one element.

Dually, we can just as easily define given a grammar notions of \emph{printer}:
\begin{definition}
  Let $G \in \Grammar\,{\mathcal A}\,X$. A \emph{printer} for $G$ is a function
  \[ (x : X) \to (w : \mathcal A^*) \times G\,w\,x \]
\end{definition}
The notion of partial printer is less useful.

If we think of an abstract grammar as a kind of constructive relation,
then that inevitably leads us to the thesis of this article:
\begin{quote}
  Grammar formalisms are logic programming languages
\end{quote}
But the most interesting point of this is that the logic that grammar
formalisms are based on is a kind of \emph{separation
logic}. Separation logic is typically used to reason about the layout
of data in memory and there is an analogous intuition for grammars: if
our memory is a string of characters in $\mathcal A$, then a grammar
describes the ways in which semantic values can be represented in
``memory'' as strings. To demonstrate this connection, we show that
the grammars assemble into a cartesian closed category that is also
monoidal biclosed, i.e., a model of the logic of bunched implications.

\section{A Category of Grammars}

Since a grammar is parameterized by a type of characters and semantic
values, there is some choice in defining a category of grammars: do we
allow these parameter types to vary or stay fixed? For modeling
grammar formalisms, it is most natural to fix a single type of
characters $\mathcal A$ and allow the types of semantic values to
vary. Intuitively this is like the typical situation of fixing a type
of tokens that are output from a lexer but defining non-terminals that
parse into many different output types.

\begin{mathpar}
  Let $\mathcal A$ be a type, the category of $\mathcal A$-grammars
  $\Grammar\,\mathcal A$ has
  \begin{enumerate}
  \item as objects...
  \item morphisms...
  \item composition/identity...
  \end{enumerate}
\end{mathpar}

To construct the various structures on this category, we observe that
it can be constructed in the following fashion: $\Set^{\mathcal
  A^*}/\Delta$, a kind of \emph{Artin Gluing} construction on toposes.

\section{Grammars as Dependent Types}

We define a 2-level dependent type theory for specifying grammar and
their output types. For the remainder of this section fix an alphabet
$\mathcal A$.

\begin{mathpar}
\item A context $\Gamma$ denotes a set.
\item A substitution $\gamma : \Delta \to \Gamma$ denotes a function
\item A type $\Gamma \vdash A$ denotes a family of sets over $\Gamma$
  and substitution $A[\gamma]$ is restriction of the family along a
  function.
\item A term $\Gamma \vdash M : A$ denotes a section of the family and
  substitution $M[\gamma]$ is restriction of a section along a
  function.
\item A grammar context $\Gamma \pipe \hat \Gamma$ denotes a grammar
  $\Grammar\,{\mathcal A^*}\,\Gamma$ and substitution
  $\hat\Gamma[\gamma]$ denotes restriction of a grammar along a
  function.
\item A grammar substitution $\Gamma\pipe \hat \gamma : \hat \Delta \to \hat \Gamma$ denotes a family of functions
  \[ \hat \gamma : (w:\mathcal A^*) \to (\gamma : \Gamma) \to \hat \Delta\,w\,\gamma \to \hat \Gamma\,w\,\gamma \]
\item A grammar type $\Gamma
  \vdash \hat A$ denotes a pair of a family of sets $\Gamma \vdash A$
  and a family of grammars $(\gamma :\Gamma) \to \Grammar\,{\mathcal
    A^*}\,(A\gamma)$.
\item A grammar term $\Gamma\pipe\hat\Gamma\vdash \hat M : \hat A$ denotes 
\end{mathpar}

\section{Subsuming Grammar Formalisms}

\end{document}